### 2023-09-03 sections/1-introduction.tex
speedup factors

Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{indyk1999sublinear}, 
clever indexing techniques such as the FM Index~\cite{simpson2010efficient}, hierarchical navigable small world (HNSW) networks [CITE], and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in large datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the 
\emph{metric entropy} and \emph{local fractal dimension}, defined in Section 2) of a dataset,
rather than its cardinality. In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
neighbors search to a hierarchical clustering approach. In this paper, we introduce CAKES (CLAM-Accelerated $K$-nearest-neighbor 
Entropy-scaling Search). CLAM (Clustered Learning of Approximate Manifolds), developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the CHESS clustering algorithm. 
Using the cluster tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform
$k$-nearest neighbors search.

>>>>

Recent approaches to tackling the exponential growth of data include locality-sensitive hashing~\cite{indyk1999sublinear}, 
clever indexing techniques such as the FM Index~\cite{simpson2010efficient}, hierarchical navigable small world (HNSW) networks~\cite{Malkov2016EfficientAR}, and entropy-scaling search~\cite{yu2015entropy, ishaq2019clustered}. 
Entropy-scaling search is a paradigm for similarity search that exploits the geometric and topological structure inherent in large datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with geometric properties (such as the 
\emph{metric entropy} and \emph{local fractal dimension}, defined in Section 2) of a dataset,
rather than its cardinality. In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search), which extended entropy-scaling $\rho$-nearest 
neighbors search to a hierarchical clustering approach. In this paper, we introduce CAKES (CLAM-Accelerated $K$-nearest-neighbor 
Entropy-scaling Search). CLAM (Clustered Learning of Approximate Manifolds), developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the CHESS clustering algorithm. 
Using the cluster tree constructed by CLAM, CAKES builds upon the capabilities of CHESS to perform
$k$-nearest neighbors search.

### 2023-09-03 sections/4-results.tex
table and knn complexity
more tables
speedup factors

Figure [ADD FIGURE NUMBER] is a violin plot which shows the performance of our algorithms
on [ADD DATASET] under [ADD DISTANCE FUNCTION] distance, using [ADD NUMBER OF SHARDS] shards.
As indicated in the title, the dataset has cardinality [ADD cardinality] and dimensionality [ADD DIMENSIONALITY].
On the horizontal axis, we have each of the $k$-NN algorithms described in Section 2.4, as well as  
an additional "algorithm" which we call "Oracle $\rho$-NN" [REPLACE WITH ACTUAL NAME].  With "Oracle $\rho$-NN", we perform $\rho$-NN search using the 
distance to the $k$-th nearest neighbor as the radius. Of course, as it is impossible to
to know the distance to the $k$-th nearest neighbor \emph{a priori}, this algorithm 
is useful only for benchmarking purposes.

The vertical axis for Figure [ADD FIGURE NUMBER] is the time query in seconds on a logarithmic scale. 
As shown in the key, different colors represent different values of $k$. 
The plot shows the distribution of query times for each algorithm for each value of $k$. 
Because of the logarithmic scale of the vertical axis, it is unsurprising that we 
see longer lower tails than upper tails for many of these algorithms.

https://github.com/URI-ABD/clam

@@

\begin{table*}[!t]
    % \renewcommand{\arraystretch}{1.15}
    \caption{Runtime performance (queries per second) of CAKES vs. other methods, $k=100$. Other methods did not report results for glove-25. Asterisk on recall value indicates that the algorithm had imperfect recall greater than or equal to 0.9995}
    \label{table:results:ann-100}
    \vskip 0.15in
    \begin{center}
    \begin{small}
    \begin{sc}
    \begin{tabular}{|l|p{1cm}|p{1cm}|}
    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{linear}} & \multicolumn{2}{|c|}{\textbf{CAKES}}  \\
    % \hline
    % deep-image             & 0.00 & 0.00                           & 0.00 & 0.00      \\
    % \hline
    % mnist                   & 0.00 & 0.00                          & 0.00 & 0.00      \\
    % \hline
    % glove-50               & 0.00 & 0.00                           & 0.00 & 0.00      \\
    % \hline 
    % glove-200              & 0.00 & 0.00                           & 0.   % \hline
    % \hline
    % mnist                   & 0.00 & 0.00                          & 0.00 & 0.00      \\
    % \hline
    % glove-50               & 0.00 & 0.00                           & 0.00 & 0.00      \\
    % \hline 
    % glove-200              & 0.00 & 0.00                           & 0.00 & 0.00      \\
    % \hline
    % lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00      \\00 & 0.00      \\
    % \hline
    % lastfm                 & 0.00 & 0.00                           & 0.00 & 0.00      \\

    \hline
    \end{tabular}
    \end{sc}
    \end{small}
    \end{center}
    \vskip -0.1in
    \end{table*}

>>>>

For $k= 10$, we report a mean $989\times$ (median $248\times$) speedup over faiss-ivf, a mean $1456\times$ (median $582\times$) speedup over bruteforce-blas, 
and a mean $63\times$ (median $49\times$) speedup over HNSW. For $k=100$, we report a mean $109\times$ (median $71\times$) speedup over faiss-ivf, a mean $382\times$ (median $267\times$) speedup over bruteforce-blas, and a mean $10\times$ (median $9\times$) speedup over HNSW.

The source code for CAKES is available under an MIT
license at https://github.com/URI-ABD/clam.

@@

\begin{table*}[!t]
    % \renewcommand{\arraystretch}{1.15}
    \caption{Runtime performance (queries per second) of CAKES vs. linear, $k=10$}
    \label{table:results:ann-alt-10}
    \vskip 0.15in
    \begin{center}
    \begin{small}
    \begin{sc}
    \begin{tabular}{|l|l|l|l|l|}
    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{linear}} & \multicolumn{2}{|c|}{\textbf{CAKES}}  \\
    &                    Recall & QPS                           & Recall & QPS                           \\
    \hline
    deep-image             & 1.000 & 0.00                           & 0.00 & 0.00      \\
    \hline
    mnist                   & 1.000 & 0.00                          & 1.000 & 0.00      \\
    \hline
    glove-50               & 1.000 & 0.00                           & 0.00 & 0.00      \\
    \hline 
    glove-200              & 1.000 & 0.00                           & 0.00 & 0.00     \\
    \hline
    lastfm                 & 1.000 & 0.00                           & 0.00 & 0.00           \\
    \hline
    \end{tabular}
    \end{sc}
    \end{small}
    \end{center}
    \vskip -0.1in
    \end{table*}

\begin{table*}[!t]
    % \renewcommand{\arraystretch}{1.15}
    \caption{Runtime performance (queries per second) of CAKES vs. linear, $k=100$}
    \label{table:results:ann-alt-100}
    \vskip 0.15in
    \begin{center}
    \begin{small}
    \begin{sc}
    \begin{tabular}{|l|l|l|l|l|}
    \textbf{Dataset}  & \multicolumn{2}{|c|}{\textbf{linear}} & \multicolumn{2}{|c|}{\textbf{CAKES}}  \\
    &                    Recall & QPS                           & Recall & QPS                           \\
    \hline
    deep-image             & 1.000 & 0.00                           & 0.00 & 0.00      \\
    \hline
    mnist                   & 1.000 & 0.00                          & 1.000 & 0.00      \\
    \hline
    glove-50               & 1.000 & 0.00                           & 0.00 & 0.00      \\
    \hline 
    glove-200              & 1.000 & 0.00                           & 0.00 & 0.00     \\
    \hline
    lastfm                 & 1.000 & 0.00                           & 0.00 & 0.00           \\
    \hline
    \end{tabular}
    \end{sc}
    \end{small}
    \end{center}
    \vskip -0.1in
    \end{table*}
